---
title: "3) Plotting model outputs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{3) Plotting model outputs}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: sentence
---

This vignette walks through using the model simulations generated in the previous vignette - "2) Simulating the model with fit parameters", to plot the figures shown in the associated publication "A mathematical model of H5N1 influenza transmission in US dairy cattle".

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(cowflu)
library(dust2)
library(monty)
library(dplyr)
library(tidyr)
```

## Load the model simulations
We load a set of model simulations from the previous vignette. Note, the results plotted in the main manuscript are from a simulation of 20,000 iterations. For simplicity, those shown below use only 500 iterations.
```{r}
# Path to the .rds file
file_path <- system.file("extdata", "example_sim_results.rds", package = "cowflu")
# Load the .rds file
Sim_results <- readRDS(file_path)
```
